---
title: "COVID_ADA Code"
author: "Amber Clover"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r GLM, include=FALSE}

glm.RR <- function(GLM.RESULT, digits = 2) {

    if (GLM.RESULT$family$family == "binomial") {
        LABEL <- "OR"
    } else if (GLM.RESULT$family$family == "poisson") {
        LABEL <- "RR"
    } else {
        stop("Not logistic or Poisson model")
    }

    COEF      <- stats::coef(GLM.RESULT)
    CONFINT   <- stats::confint(GLM.RESULT)
    TABLE     <- cbind(coef=COEF, CONFINT)
    TABLE.EXP <- round(exp(TABLE), digits)

    colnames(TABLE.EXP)[1] <- LABEL

    TABLE.EXP
}
```


```{r Data Import, include=FALSE}

#Data import
covid <- read.csv("/Users/amberclover/Desktop/stateleveldata.csv")
View(covid)

#relevel Region to make the South the reference group
covid <- within(covid, Region <-relevel(Region, ref = "West"))

# Distribution Shape - Cases
d <- density(covid$cases)
plot(d, xlim=c(1,200000))
hist(covid$case,breaks=c(0,2500,5000,7500,10000,12500,15000,17500,20000,22500, 25000, 27500, 30000, 35000, 40000, 45000, 50000, 200000))

# Distribution Shape - Deaths
d <- density(covid$deaths)
plot(d, xlim=c(1,20000))
hist(covid$deaths,breaks=c(0,1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000))

# Packages
library(MASS)
library(lmtest)
library(sandwich)

```

```{r Model without Offset, echo=FALSE, warning=FALSE, message=FALSE}
# base model of all states - cases by pop. density
model.0 <- glm(cases ~ Region, family ="poisson", data = covid)

summary(model.0)

# base model of all states - death by pop. density
model.2 <- glm(deaths ~ Region, family ="poisson", data = covid)

summary(model.2)
```

```{r Poisson Model for Cases, echo=FALSE, warning=FALSE, message=FALSE}

############### model including covariates with South as the reference group ###############

model.1 <- glm(cases ~ Region + Pop.Density.Cat + offset(log(population.Time.as.Risk)), family ="poisson", data = covid)

summary(model.1)

# IRRs
glm.RR(model.1, 2)

```





```{r Poisson Model for Deaths, echo=FALSE, warning=FALSE, message=FALSE}

############### model including covariates with South as the reference group ###############

model.11 <- glm(deaths ~ Region + Pop.Density.Cat + offset(log(population.Time.as.Risk)), family ="poisson", data = covid)

summary(model.11)

# IRRs
glm.RR(model.11, 2)

```


```{r Standard Errors for Case Model, echo=FALSE, warning=FALSE, message=FALSE}

######### Poisson model cases with SE estimated via robust variance estimator #########
library(sandwich)
library(lmtest)
coeftest(model.1, vcov = sandwich)

cov.model <- vcovHC(model.1, type="HC0") 
std.err <- sqrt(diag(cov.model)) #robust std erros for coefficients

#summary table of estumates
r.est <- cbind(Estimate= coef(model.1), "Robust SE" = std.err,
"Pr(>|z|)" = 2 * pnorm(abs(coef(model.1)/std.err), lower.tail=FALSE),
LL = coef(model.1) - 1.96 * std.err,
UL = coef(model.1) + 1.96 * std.err) 

r.est

r.est1 <- cbind(Estimate= exp(coef(model.1)), LL = exp(coef(model.1) - 1.96 * std.err),
UL = exp(coef(model.1) + 1.96 * std.err)) 
r.est1

```



```{r Standard Errors Death Model, echo=FALSE, warning=FALSE, message=FALSE}

######### Poisson model deaths with SE estimated via robust variance estimator #########
coeftest(model.11, vcov = sandwich)

cov.model.11 <- vcovHC(model.11, type="HC0") #type specifies variance estimator method, the vcovHC function gives the covariance matrix of the coefficient estimates. Need this to estimate robust SE
std.err1 <- sqrt(diag(cov.model.11)) #estimate robust standard error for each coefficient

#make a summary table of coefficients, robust SEs, and LL and UL confidence intervals (log scale)
r.est2 <- cbind(Estimate= coef(model.11), "Robust SE" = std.err1,
"Pr(>|z|)" = 2 * pnorm(abs(coef(model.11)/std.err1), lower.tail=FALSE),
LL1 = coef(model.11) - 1.96 * std.err1,
UL1 = coef(model.11) + 1.96 * std.err1) 

r.est2

r.est3<-cbind(Estimate= exp(coef(model.11)), LL = exp(coef(model.11) - 1.96 * std.err1),
UL = exp(coef(model.11) + 1.96 * std.err1)) 
r.est3
```


```{r Neg Bin Models, echo=FALSE, warning=FALSE, message=FALSE}
#neg bin model cases to check for overdispersion
model.1nb <- glm.nb(cases ~ Region + Pop.Density.Cat + offset(log(population.Time.as.Risk)), data = covid)
summary(model.1nb)

#neg bin model deaths to check for overdispersion
model.11nb <- glm.nb(deaths ~ Region + Pop.Density.Cat + offset(log(population.Time.as.Risk)), data = covid)
summary(model.11nb)
```


```{r compare models, echo=FALSE, warning=FALSE, message=FALSE}
#compare case models
library(lmtest)
lrtest(model.1, model.1nb)

#compare death models
lrtest(model.11, model.11nb)
```

*Note: The negative binomial models for cases (`X^2`=229759, p<.05) significantly improved the fit of both models and should be used. The negative binomial models for deaths (`X^2`=14355, p<.05) significantly improved the fit of both models and should be used.*


```{r stargazer, echo=FALSE, warning=FALSE, message=FALSE}
#Cases - Poisson v. Neg Bin
library(stargazer)
stargazer(model.1, model.1nb, title="Case Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)

#Deaths - Poisson v. Neg Bin
stargazer(model.11, model.11nb, title="Death Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)
```

*Interpretation: SE from negative binomial models were much larger than those of the Poisson for both the case and death models, suggesting the presence of overdispersion.*

```{r, echo=FALSE, warning=FALSE, message=FALSE}

## Negative Binomial model with SE estimated via robust variance estimator for cases
robust1 <- coeftest(model.1nb, vcov = sandwich)

#print results
robust1

## Negative Binomial model with SE estimated via robust variance estimator for deaths
robust11 <- coeftest(model.11nb, vcov = sandwich)

#print results
robust11

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Neg Bin IRRs for Cases
est.cases <- cbind(IRR = coef(model.1nb), "2.5%"=robust1[,1]-1.96*robust1[,2], 
             "97.5%"=robust1[,1]+1.96*robust1[,2])
exp(est.cases)

#Neg Bin IRRs for Deaths
est.deaths <- cbind(IRR = coef(model.11nb), "2.5%"=robust11[,1]-1.96*robust11[,2], 
             "97.5%"=robust11[,1]+1.96*robust11[,2])
exp(est.deaths)

```




